The classical mutual information in mean-field spin glass models

### 2015-06-01 sk_model_v15.tex
latest

We investigate the \emph{classical} R\'enyi entropy $S_n$ and the associated mutual 
information ${\mathcal I}_n$ in the Sherrington-Kirkpatrick (S-K) model, which is the 
paradigm model of mean-field spin glasses. 
Using both classical Monte Carlo simulation and analytical tools we investigate the S-K model 
on a booklet of $n$-sheets.
This is obtained by gluing together $n$ independent copies of the model, and it is the 
main ingredient to construct the R\'enyi entaglement-related quantities. We find that at low temperature 
the S-K model is in a glassy phase, whereas at high temperature it exhibits paramagnetic 
behavior. Interestingly, the critical temperature of the paramagnetic-glassy transition 
depends in a non-trivial way on the geometry of the booklet. At high-temperatures we provide 
the exact solution of the model exploiting the replica symmetry. This is the permutation 
symmetry among the fictitious replicas that are used in the replica trick to perform disorder 
averages. In the low-temperature glassy phase the replica symmetry has 
to be broken. Using a generalization of the celebrated Parisi solution, we provide 
approximated results for thermodynamic quantities and for $S_n$ and ${\mathcal I}_n$. 
Finally, we show that both $S_n$ and ${\mathcal I}_n$ exhibit a volume law in the whole phase 
diagram. Their densities $S_n/N$ and ${\mathcal I}_n/N$ are finite and smooth 
function of the temperature, in the thermodynamic limit. In contrast with local spin models, 
${\mathcal I}_n/N$ does not exhibit any crossing for different system sizes at the 
critical point.

@@

Besides being ubiquitous in nature, disorder leads to several intriguing 
physical phenomena. 
Arguably, \emph{spin glasses} represent one of the most prototypical examples
of disorder.
While at any finite temperature disorder prevents the 
usual magnetic ordering, at low-enough temperatures these systems display a 
new and exotic type of ``order''. In the past decades an intense theoretical 
effort has been devoted to characterizing this spin glass order, the nature of 
the paramagnetic-glassy transition, and that of the associated order 
parameter~\cite{binder-1986,parisi-book,young-1998,nishimori-book,castellani-2005}.

All these issues have been thoroughly addressed in the Sherrington-Kirkpatrick 
(S-K) model~\cite{sherrington-1978,sherrington-1978-prl}, which is exactly 
solvable. The S-K model is a \emph{classical} Ising model on the fully-connected 
graph of $N$ sites, with quenched random interactions. Its hamiltonian reads  
%
\begin{equation}
{\mathcal H}=-\sum\limits_{1\le i<j\le N}J_{ij}S_i S_j-
h\sum\limits_{1\le i\le N}S_i.
\label{SK-intro}
\end{equation}
%
Here $S_i=\pm 1$ are classical Ising spins, $h$ is an external magnetic field, and 
$J_{ij}$ are uncorrelated (from site to site) random variables. The S-K model hosts 
a low-temperature glassy phase, which is separated from a high-temperature paramagnetic 
one by a second order phase transition. In contrast with the mean-field nature of 
Eq.~\eqref{SK-intro}, the solution of the S-K model has been a mathematical challenge. 
Although it was proposed as an ansatz by Parisi~\cite{parisi-1980} more than thirthy 
years ago, its rigourous proof was obtained only recently~\cite{talagrand-2006}. Moreover, 
the solution exhibits several intricate features, such as lack of self-averaging~\cite{pastur-1991}, 
ultrametricity~\cite{mezard-1984,rammal-1986}, and replica symmetry breaking~\cite{parisi-book,
castellani-2005}. The last refers to the breaking of the permutation symmetry among the 
fictitious replicas of the system, which in the standard replica trick~\cite{cardy-book} are 
used to perform disorder averages. Finally, although the applicability of the S-K 
model to describe realistic spin glasses is still highly debated~\cite{yucesoy-2012,
billoire-2012,yucesoy-2013}, we should mention that there are recent proposals on how to 
realize this model in cold-atomic gases~\cite{morrison-2008,rotondo-2015}, or in laser 
systems~\cite{ghofraniha-2015}.

In the last decade entanglement-related quantities have emerged as valuable tools to 
understand the physics of complex systems~\cite{amico-2008,eisert-2009,calabrese-2009,
cc-rev}, both classical and quantum. For instance, at a conformally invariant critical 
point entanglement measures contain universal information about the underlying 
conformal field theory (CFT), such as the central charge~\cite{holzhey-1994,vidal-2003,
calabrese-2004,calabrese-2012}. For classical spin models a lot of attention has been 
focused on the \emph{classical} R\'enyi entropy~\cite{jaconis-2013,stephan-2014}. Given 
a bipartition of the system into two complementary subregions $A$ and $B$, the classical 
R\'enyi entropy $S_n(A)$ (with $n\in\mathbb{N}$) is defined as 
%
\begin{equation}
S_n(A)\equiv \frac{1}{1-n}\log\Big(\sum\limits_{i_A\in{\mathcal C}_A} p^n_{i_A}
\Big)
\label{renyi-intro}
\end{equation}
%
Here ${\mathcal C}_A$ denotes the set of all the possible spin configurations in part 
$A$, whereas $p_{i_A}$ is the probability of the configuration $i_A$. 
%Clearly, $p_{i_A}\equiv\sum_{j_b\in{\mathcal C}_B}p_{i_A,j_B}$, with $p_{i_A,j_B}$ 
%being the probability that the system is in the configuration $i_A\in{\mathcal C}_A$
%in part $A$, and $j_B\in{\mathcal C}_B$ in part $B$. For a spin model at inverse 
%temperature $\beta$, one has $p_{i_A,j_B}\equiv e^{-\beta E(i_A,j_B)}/Z(\beta)$, 
%where $E(i_A,j_B)$ is the energy of the system, and $Z(\beta)$ its partition function. 
Alternatively, $S_n(A)$ can be obtained from the partition function of the model on 
an \emph{ad hoc} defined ``booklet'' geometry (see section~\ref{booklet} for its 
precise definition). This consists of $n$ independent and identical replicas (the 
booklet ``sheets'') of the model. Notice that these \emph{physical} replicas are different 
from the \emph{fictitious} ones used to perform the disorder average. Each sheet is 
divided into two parts $A$ and $B$, containing $N_A$ and $N_B$ spins, respectively. 
The spins in part $A$ of different sheets are constrained to be equal. It is 
convenient to introduce the aspect ratio $0\le\omega\le1$ as  
%
\begin{equation}
\label{a-ratio}
\omega\equiv \frac{N_A}{N}.
\end{equation}
%
Clearly,  Eq.~\eqref{renyi-intro} can be extended to quantum systems~\cite{stephan-2009,
stephan-2010,stephan-2011,zaletel-2011,alcaraz-2013,stephan-2014-a} by interpreting 
$p_{i_A}$ as the probability of finding part $A$ in the \emph{quantum} configuration 
$i_A$. From $S_n(A)$, one defines the classical mutual information ${\mathcal I}_n(A,B)$ 
as 
%
\begin{equation}
{\mathcal I}_n(A,B)\equiv S_n(A)+S_n(B)-S_n(A\cup B). 
\end{equation}
%
For local models ${\mathcal I}_n$ obeys the area law~\cite{wolf-2008} ${\mathcal I}_n(A)
\propto\ell$, with $\ell$ being the length of the boundary between $A$ and $B$. Remarkably, 
for different $\ell$, the ratio ${\mathcal I}_n/\ell$ exhibits a crossing at a second 
order phase transition~\cite{jaconis-2013}, and it can be used as a diagnostic tool for 
critical behaviors. For conformally invariant critical models more universal information 
can be extracted from the subleading contributions, i.e., the corrections to the area 
law, of ${\mathcal I}_n$~\cite{stephan-2014}.

Recently, the study of the interplay between disorder and entanglement became a fruitful 
research area~\cite{refael-2009}. However, the behavior of entanglement-related quantities 
in glassy phases, and at glassy critical points, has not been explored yet. Here we investigate 
both the classical R\'enyi entropy $S_n(A)$ and the mutual information ${\mathcal I}_n$ 
in the S-K model, using classical Monte Carlo simulations and analytical tools. We often 
restrict ourselves to the case with $n=2$, as this is where numerical simulation is most
efficient . As usual in disordered system, we 
focus on disorder-averaged quantities, considering $[S_n]$ and $[{\mathcal I}_n]$, 
with the brackets $[\cdot]$ denoting the average over different realizations of 
$J_{ij}$ in Eq.~\eqref{SK-intro}.

We start discussing the thermodynamic phase diagram of the S-K model on the $n$-sheets 
booklet, as a function of temperature, and the aspect ratio $\omega$ (cf. Eq.~\eqref{a-ratio}). 
We find that for any $\omega$ the S-K model exhibits a low-temperature glassy phase, which 
is divided by the standard paramagnetic one at high temperatures by a critical point. 
Surprisingly, the critical temperature exhibits a non-trivial dependence on $\omega$ 
that we are able to determine analytically. 
In the high-temperature region, for any fixed $\omega$, the permutational symmetry 
among the replicas, both the physical and the fictitious ones, is preserved. This 
allows us to provide an exact analytic expression for the free energy of the model 
and several derived quantities, such as the internal energy. We check our result 
against Monte Carlo simulations, finding perfect agreement. Oppositely, we provide 
both numerical and analytical evidence that in the low-temperature phase, for any 
$\omega$, the replica symmetry has to be broken. Specifically, we numerically observe 
that the replica-symmetric (RS) result for the internal energy  is systematically 
lower than the Monte Carlo data, as in the standard S-K model~\cite{sherrington-1978-prl,
sherrington-1978}. This discrepancy becomes larger upon lowering the temperature.

@@

Inspired by the Parisi scheme~\cite{parisi-1979}, we devise a systematic 
way of breaking the replica symmetry in successive steps. Precisely, our scheme  
breaks only the symmetry among the fictitious replicas, preserving that among the 
physical ones. Although this appears natural, we are not able to provide a rigourous 
proof that this is the correct symmetry breaking pattern. In particular, we cannot 
exclude that the symmetry among the physical replicas has to be broken. As a consequence, 
our scheme should be regarded as an approximation, and not as an exact solution. 
Finally, we should mention that we restrict ourselves to the one-level replica symmetry 
breaking ($1$-RSB). This, however, is a minor limitation because, in principle, it is 
possible to implement in our ansatz the full breaking of the replica symmetry 
($\infty$-RSB). Surprisingly, we numerically observe that the $1$-RSB approximation 
for the internal energy is in excellent agreement with the Monte Carlo data for  
$\beta\lesssim 3$, whereas the RS approximation fails already at $\beta\approx 1$. 
This suggests that the $1$-RSB ansatz captures correctly some aspects of the 
replica symmetry breaking.

In this scenario the effects of the replica symmetry breaking are 
reflected in the behavior of $[S_2]$ and $[{\mathcal I}_2]$. First, since $[S_2]$ exhibits 
the volume law behavior $[S_2]\propto N$, we consider its density $[S_2]/N$. For finite-size 
systems, and for any $\omega$, we find that $[S_2]/N$ exhibits a maximum at infinite 
temperature, and it is a monotonically decreasing function of the temperature. More 
specifically, the Monte Carlo data suggest a vanishing behavior in the zero-temperature 
limit. Moreover, the finite-size corrections are neglegible at high temperatures, 
whereas they increase upon lowering the temperature. A standard finite-size scaling 
analysis shows that in the thermodynamic limit $[S_2]/N$ is finite and smooth as a 
function of $\beta$ and $\omega$, in the whole phase diagram. In the paramagnetic phase 
we are able to calculate the functional form of $[S_2]$, using the replica-symmetric 
approximation. This perfectly matches the Monte Carlo data. At low temperatures deviations 
from the replica-symmetric result are observed, reflecting the replica symmetry breaking. 
Remarkably, we observe that the one-step replica symmetry breaking ($1$-RSB) approximation 
$S_2^{1\textrm{-}RSB}/N$ is in full agreement with the Monte Carlo data for $\beta
\lesssim 3$, confirming what is observed for the internal energy.

Finally, we consider the R\'enyi mutual information $[{\mathcal I}_2]$. This  
obeys a volume law for any $\beta$ and $\omega$, in sharp contrast with local spin models, 
where an area law is observed~\cite{wolf-2008}. The mutual information density $[{\mathcal I}_2]/N$ 
vanishes in both the infinite-temperature and the zero-temperature limits, for any $\omega$. 
Surprisingly, for different sizes ${\mathcal I}_2/N$ does not exhibit any crossing at the 
paramagnetic-glassy transition, in striking contrast with local spin models~\cite{jaconis-2013}. 
In the thermodynamic limit $[{\mathcal I}_2]/N$ is finite and smooth (as a function of $\beta$) 
in the whole phase diagram. Interestingly, for any $\omega$, ${\mathcal I}_2/N$ exhibits a 
maximum for $\beta\approx 1$. However, the position of this maximum is not simply related to 
paramagnetic-glassy transition. Finally, at high temperature $[{\mathcal I}_2]/N$ is fully 
described analytically within the replica symmetric approximation, whereas at low temperatures 
one has to include  the effects of the replica symmetry breaking. Similar to $[S_2]$, we observe 
that the $1$-RSB approximation for $[{\mathcal I})_2]/N$, is in good agreement with the Monte Carlo 
data for $\beta \lesssim 3$.

The Article is organized as follows. In section~\ref{booklet} we introduce the classical 
R\'enyi entropy and the mutual information, reviewing their representation in terms of 
the booklet partition functions. In section~\ref{the-model} we present the structure of 
the  the solution of the S-K model on the booklet. Specifically, we discuss the replica 
trick, which is used to perform disorder averages, and the saddle point approximation in 
the thermodynamic limit . Section~\ref{solution} is concerned with the replica-symmetric 
(RS) approximation. In subsection~\ref{para-section} we focus on the high-temperature 
region where this approximation becomes exact. In subsection~\ref{rs-section} we discuss 
the RS ansatz in the low-temperature region. Section~\ref{rsb-1-section} is devoted to   
presenting the one-step replica symmetry breaking ($1$-RSB) approximation. In 
section~\ref{mc-results} we check the validity of both the RS and the $1$-RSB approximations 
against Monte Carlo data. Section~\ref{Renyi-section} and~\ref{I2-section} present our results 
for the classical R\'enyi entropy and the mutual information, respectively. Finally, 
we conclude in section~\ref{conclusions}.

@@

Given a generic \emph{classical} spin model at inverse temperature $\beta$, 
the R\'enyi mutual information ${\mathcal I}_n(A,B,\beta)$ (with $n\in\mathbb{N}$
and $A \cup B$ is the entire system)) 
is defined as~\cite{jaconis-2013,stephan-2014} 
%
\begin{equation}
\label{MI}
{\mathcal I}_n(A,B,\beta)\equiv\frac{1}{1-n}\log\Big(
\frac{Z(A,n,\beta)Z(B,n,\beta)}{Z(\beta)^nZ(n\beta)}
\Big).
\end{equation}
%
Here $Z(A,n,\beta)$ is the partition function of the model on the so-called 
$n$-sheet booklet, whereas $Z(\beta)$ and $Z(n\beta)$ are the partition functions 
on a the plane (a single sheet) at inverse temperatures $\beta$ and $n\beta$, 
respectively. The booklet geometry is illustrated in Figure~\ref{cartoon}. The 
sheet is divided into two parts $A$ and $B$ (cf. Figure~\ref{cartoon} (a)). 
Then $n$ copies of the system (``sheets'') are ``glued'' together to form the 
booklet. The spins in part $A$ of different sheets are identified (cf. 
Figure~\ref{cartoon} (b)). Notice that a similar construction~\cite{guerra-2002} 
plays an important role in the mathematical proof of the Parisi ansats. While spins 
in $B$ are at inverse temperature $\beta$, the ones in $A$ are at the effective 
temperature $n\beta$. The definition of $Z(B,n,\beta)$ is obtained by exchanging 
$A$ and $B$. Alternatively, ${\mathcal I}_n$ can be rewritten as 
%
\begin{equation}
{\mathcal I}_n(A,B,\beta)=S_n(A)+S_n(B)-S_n(A\cup B),
\end{equation}
%
where $S_n(X)$ is the ``booklet'' classical R\'enyi entropy
%
\begin{equation}
\label{renyi}
S_n(X)\equiv \frac{1}{1-n}\log\left(\frac{Z(X,n,\beta)}{Z^n(\beta)}\right).
\end{equation}
%
Notice that $S_n(X)$ corresponds to $S_n(X)\equiv1/(n-1)\log(F(X,n,\beta))
-n\log(F(\beta))$, where $F(X,n,\beta)\equiv\log(Z(X,n\beta))$ and $F(\beta)
\equiv\log(Z(\beta))$ are related (apart from a factor $\beta$) to the 
free energy of the model on the booklet and on the plane, respectively.
Since here we focus on the effects of the disorder, we consider 
the disorder-averaged mutual information $[{\mathcal I}_n]$ 
%
\begin{equation}
[{\mathcal I}_n(A,B,\beta)]=[S_n(A)]+[S_n(B)]-[S_n(A\cup B)],
\end{equation}
%
where the square brackets $[\cdot]$ denote the average over different disorder 
realizations (cf. section~\ref{the-model}). Notice that $[{\mathcal I}_n]$ is 
directly obtained from the so-called quenched averaged free energy $[F(X,n,\beta)]$, 
which is usually considered in disordered systems~\cite{cardy-book}.

For clean (i.e., without disorder) \emph{local} spin models ${\mathcal I}_n$ 
obeys the boundary law 
%
\begin{equation}
{\mathcal I}_n(A,B,\beta)=\alpha_n\ell+{\mathcal G}_n+\gamma_n,
\label{GMI}
\end{equation}
%
with $\ell$ the length of the boundary between $A$ and $B$, and 
$\alpha_n,\gamma_n$ two non-universal constants. Here ${\mathcal G}_n$ 
is the so-called geometric mutual information~\cite{stephan-2014}. 
Interestingly, for critical systems ${\mathcal G}_n$ depends only on the 
geometry of $A$ and $B$, and it is universal. In particular, for conformally 
invariant models ${\mathcal G}_n$ can be calculated using standard methods of 
conformal field theory (CFT), and it allows to numerically 
extract universal information about the CFT, such as the central 
charge~\cite{stephan-2014}.  Remarkably, there is numerical evidence that  
${\mathcal I}_n/\ell$ exhibits a crossing for different 
system sizes at a second order phase transition, implying that it 
can be used as a diagnostic tool to detect phase 
transitions~\cite{jaconis-2013}.

@@

In this section we present the solution of the S-K model on the booklet, using 
the replica symmetric approximation. In subsection~\ref{para-section} we focus 
on the paramagnetic phase at high temperature, where this approximation is exact. 
In particular, we show that the behavior of the model is fully determined by the 
diagonal part of the overlap tensor $q_{\gamma\gamma'}^{rr'}$. The general structure 
of $q_{\gamma\gamma'}^{rr'}$ in the replica-symmetric (RS) ansatz  is discussed in 
subsection~\ref{rs-section}. This allows us to determine the critical temperature of 
the paramagnetic-glassy transition. Similar to the standard S-K model, the RS ansatz 
is not correct in the low-temperature region.

>>>>

We investigate the \emph{classical} R\'enyi entropy $S_n$ and the associated mutual 
information ${\mathcal I}_n$ in the Sherrington-Kirkpatrick (S-K) model, which is the 
paradigm model of mean-field spin glasses. 
Using both classical Monte Carlo simulation and analytical tools we investigate the 
S-K model on a booklet of $n$-sheets.
This is obtained by gluing together $n$ independent copies of the model, and it is the 
main ingredient to construct the R\'enyi entaglement-related quantities. We find that at 
low temperature the S-K model is in a glassy phase, whereas at high temperature it exhibits 
paramagnetic behavior. Interestingly, the temperature of the paramagnetic-glassy 
transition depends in a non-trivial way on the geometry of the booklet. At high-temperatures 
we provide the exact solution of the model exploiting the replica symmetry. This is the 
permutation symmetry among the fictitious replicas that are used in the replica trick to 
perform disorder averages. In the glassy phase the replica symmetry has to be broken. 
Using a generalization of the celebrated Parisi solution, we provide approximated results 
for $S_n$ and ${\mathcal I}_n$, and for standard thermodynamic quantities. Both $S_n$ 
and ${\mathcal I}_n$ exhibit a volume law in the whole phase diagram. We characterize 
analytically and numerically the behavior of the corresponding densities $S_n/N$ and 
${\mathcal I}_n/N$, in the thermodynamic limit. Remarkably, at the critical point 
${\mathcal I}_n/N$ does not exhibit any crossing for different system sizes, in 
contrast with local spin models.

@@

Besides being ubiquitous in nature, disorder leads to several intriguing 
physical phenomena. 
Arguably, \emph{spin glasses} represent one of the most prototypical examples
of exotic behavior induced by disorder.
While at any finite temperature disorder prevents the 
usual magnetic ordering, at low-enough temperatures these systems display a 
new type of ``order''. In the past decades an intense theoretical 
effort has been devoted to characterizing this spin glass order, the nature of 
the paramagnetic-glassy transition, and that of the associated order 
parameter~\cite{binder-1986,parisi-book,young-1998,nishimori-book,castellani-2005}.

All these issues have been thoroughly addressed in the Sherrington-Kirkpatrick 
(S-K) model~\cite{sherrington-1978,sherrington-1978-prl}, which is exactly 
solvable. The S-K model is a \emph{classical} Ising model on the fully-connected 
graph of $N$ sites, with quenched random interactions. Its hamiltonian reads  
%
\begin{equation}
{\mathcal H}=-\sum\limits_{1\le i<j\le N}J_{ij}S_i S_j-
h\sum\limits_{1\le i\le N}S_i.
\label{SK-intro}
\end{equation}
%
Here $S_i=\pm 1$ are classical Ising spins, $h$ is an external magnetic field, and 
$J_{ij}$ are uncorrelated (from site to site) random variables. The S-K model hosts 
a low-temperature glassy phase, which is separated from a high-temperature paramagnetic 
one by a second order phase transition. In contrast with its mean-field nature, the 
solution of the S-K model has been a mathematical challenge. 
Although it was proposed as an ansatz by Parisi~\cite{parisi-1980} more than thirthy 
years ago, its rigourous proof was obtained only recently~\cite{talagrand-2006}. Moreover, 
the solution exhibits several intricate features, such as lack of self-averaging~\cite{pastur-1991}, 
ultrametricity~\cite{mezard-1984,rammal-1986}, and replica symmetry breaking~\cite{parisi-book,
castellani-2005}. The last refers to the breaking of the permutation symmetry among the 
fictitious replicas of the system, which in the standard replica trick~\cite{cardy-book} are 
used to perform disorder averages. Finally, although the applicability of the S-K 
model to describe realistic spin glasses is still highly debated~\cite{yucesoy-2012,
billoire-2012,yucesoy-2013}, we should mention that there are recent proposals on how to 
realize this model in cold-atomic gases~\cite{morrison-2008,rotondo-2015}, or in laser 
systems~\cite{ghofraniha-2015}.

In the last decade entanglement-related quantities have emerged as valuable tools to 
understand the physics of complex systems~\cite{amico-2008,eisert-2009,calabrese-2009,
cc-rev}, both classical and quantum. For instance, at a conformally invariant critical 
point entanglement measures contain universal information about the underlying 
conformal field theory (CFT), such as the central charge~\cite{holzhey-1994,vidal-2003,
calabrese-2004,calabrese-2012}. For classical spin models a lot of attention has been 
focused on the \emph{classical} R\'enyi entropy~\cite{jaconis-2013,stephan-2014}. Given 
a bipartition of the system into two complementary subregions $A$ and $B$, the classical 
R\'enyi entropy $S_n(A)$ (with $n\in\mathbb{N}$) is defined as 
%
\begin{equation}
S_n(A)\equiv \frac{1}{1-n}\log\Big(\sum\limits_{i_A\in{\mathcal C}_A} p^n_{i_A}
\Big)
\label{renyi-intro}
\end{equation}
%
Here ${\mathcal C}_A$ denotes the set of all the possible spin configurations in part 
$A$, whereas $p_{i_A}$ is the probability of the configuration $i_A$. 
Alternatively, $S_n(A)$ can be obtained from the partition function of the model on 
an \emph{ad hoc} defined ``booklet'' geometry (see section~\ref{booklet} for its 
precise definition). This consists of $n$ independent and identical replicas (the 
booklet ``sheets'') of the model. Notice that these \emph{physical} replicas are different 
from the \emph{fictitious} ones used to perform the disorder average. Each sheet is 
divided into two parts $A$ and $B$, containing $N_A$ and $N_B$ spins, respectively. 
The spins in part $A$ of different sheets are constrained to be equal. It is 
convenient to introduce the aspect ratio $0\le\omega\le1$ as  
%
\begin{equation}
\label{a-ratio}
\omega\equiv \frac{N_A}{N}.
\end{equation}
%
Eq.~\eqref{renyi-intro} can be used for quantum systems by interpreting $p_{i_A}$ as 
the probability of finding part $A$ in the \emph{quantum} configuration $i_A$. In particular, 
for $n=1$ it defines the subsystem Shannon entropy~\cite{alcaraz-2013,stephan-2014-a}. 
From $S_n(A)$, one defines the classical mutual information ${\mathcal I}_n(A,B)$ 
as 
%
\begin{equation}
{\mathcal I}_n(A,B)\equiv S_n(A)+S_n(B)-S_n(A\cup B). 
\end{equation}
%
For local models ${\mathcal I}_n$ obeys the area law~\cite{wolf-2008} ${\mathcal I}_n(A)
\propto\ell$, with $\ell$ the length of the boundary between $A$ and $B$. Remarkably, 
for different $\ell$, the ratio ${\mathcal I}_n/\ell$ exhibits a crossing at a second 
order phase transition~\cite{jaconis-2013}, implying that it can be used as a diagnostic 
tool for critical behaviors. For conformally invariant critical models more universal 
information can be extracted from the area-law corrections of ${\mathcal I}_n$~\cite{stephan-2014}.

Recently, the study of the interplay between disorder and entanglement became a fruitful 
research area~\cite{refael-2009}. However, the behavior of entanglement-related quantities 
in glassy phases, and at glassy critical points, has not been explored yet (see 
Ref.~\onlinecite{castelnovo-2010} for some interesting results). Here we investigate 
both the classical R\'enyi entropy $S_n(A)$ and the mutual information ${\mathcal I}_n$ 
in the S-K model, using classical Monte Carlo simulations and analytical tools. We often 
restrict ourselves to the case with $n=2$, as this is where numerical simulations are 
most efficient. As usual in disordered system, we focus on disorder-averaged quantities, 
considering $[S_n]$ and $[{\mathcal I}_n]$, with the brackets $[\cdot]$ denoting the 
average over different realizations of $J_{ij}$ (cf. Eq.~\eqref{SK-intro}).

We start discussing the thermodynamic phase diagram of the S-K model on the $n$-sheets 
booklet, as a function of temperature, and the aspect ratio $\omega$ (cf. Eq.~\eqref{a-ratio}). 
We show that for any fixed $\omega$ the S-K model exhibits a low-temperature glassy phase, 
which is divided by the standard paramagnetic one at high temperatures by a critical point. 
Surprisingly, the critical temperature exhibits a non-trivial dependence on $\omega$ 
that we are able to determine analytically. 
In the high-temperature region the permutational symmetry among the replicas, both the 
physical and the fictitious ones, is preserved. This allows us to provide an exact analytic 
expression for the free energy of the model and several derived quantities, such as the internal 
energy. We compare our results with Monte Carlo simulations, finding perfect agreement. 
Oppositely, we provide evidence that in the low-temperature phase the replica symmetry is  
broken. Precisely, we numerically observe that the replica-symmetric (RS) result for the 
internal energy  is systematically lower than the Monte Carlo data, as in the standard 
S-K model~\cite{sherrington-1978-prl,sherrington-1978}. This discrepancy becomes larger 
upon lowering the temperature.

@@

Inspired by the Parisi scheme~\cite{parisi-1979}, we devise a systematic way of breaking 
the replica symmetry in successive steps. In our scheme we break only the symmetry among 
the fictitious replicas, preserving that among the physical ones. Although this appears 
natural, we are not able to provide a rigourous proof that this is the correct symmetry 
breaking pattern. In particular, we cannot exclude that the symmetry among the physical 
replicas has to be broken. As a consequence, our scheme should be regarded as an 
approximation, and not as an exact solution. Finally, we restrict ourselves to the 
one-level replica symmetry breaking ($1$-RSB), although, in principle, it is possible 
to treat within our scheme the full breaking of the replica symmetry ($\infty$-RSB). 
Surprisingly, we observe that the $1$-RSB approximation for the internal energy is in 
excellent agreement with the Monte Carlo data for  $\beta\lesssim 3$, whereas the RS 
approximation fails already at $\beta\approx 1$. This suggests that the $1$-RSB ansatz 
captures correctly some aspects of the replica symmetry breaking.

Clear signatures of the replica symmetry breaking are observed in the behavior of $[S_2]$. 
First, since $[S_2]$ exhibits the volume law behavior $[S_2]\propto N$, 
we consider its density $[S_2]/N$. For finite-size systems, and for any $\omega$, $[S_2]/N$ 
exhibits a maximum at infinite temperature, and it is a decreasing function of the temperature, 
as expected. Finite-size corrections are neglegible at high temperatures, whereas they increase 
upon lowering the temperature. In the paramagnetic phase we are able to determine the functional 
form of $[S_2]/N$ in the thermodynamic limit, using the replica-symmetric approximation. This 
perfectly matches the Monte Carlo data. At low temperatures deviations from the RS result are 
observed, reflecting the replica symmetry breaking. Remarkably, we observe that the one-step 
replica symmetry breaking ($1$-RSB) result $S_2^{1\textrm{-}RSB}/N$ is in full agreement with 
the Monte Carlo data for $\beta\lesssim 3$, confirming what is observed for the internal 
energy.

Finally, we consider the R\'enyi mutual information $[{\mathcal I}_2]$. This obeys a volume 
law for any $\beta$ and $\omega$, in contrast with local spin models, where an area law is 
observed~\cite{wolf-2008}. The corresponding density $[{\mathcal I}_2]/N$ vanishes in 
both the infinite-temperature and the zero-temperature limits. Surprisingly, ${\mathcal I}_2/N$ 
does not exhibit any crossing for different system sizes at the paramagnetic-glassy transition.
This is in striking contrast with local spin models~\cite{jaconis-2013}. Interestingly, for 
any $\omega$, ${\mathcal I}_2/N$ exhibits a maximum for $\beta\approx 1$. However, the 
position of this maximum is not simply related to the paramagnetic-glassy transition. Finally, 
at high temperature $[{\mathcal I}_2]/N$ is fully described analytically by the RS result 
$[{\mathcal I}_2]^{RS}/N$, whereas at low temperatures one has to include the effects of the 
replica symmetry breaking. Similar to $[S_2]$, we observe that the $1$-RSB approximation 
$[{\mathcal I}^{1\textrm{-}RSB}_2]/N$, is in good agreement with the Monte Carlo data 
for $\beta\lesssim 3$.

The Article is organized as follows. In section~\ref{booklet} we introduce the classical 
R\'enyi entropy and the mutual information, reviewing their representation in terms of 
the booklet partition functions. In section~\ref{the-model} we present the structure of 
the  the solution of the S-K model on the booklet. Specifically, we discuss the replica 
trick, which is used to perform disorder averages, and the saddle point approximation in 
the thermodynamic limit . Section~\ref{solution} is concerned with the RS approximation. 
In subsection~\ref{para-section} we focus on the high-temperature 
region, where this approximation becomes exact. In subsection~\ref{rs-section} we discuss 
the RS ansatz in the low-temperature region. Section~\ref{rsb-1-section} is devoted to   
the $1$-RSB approximation. In section~\ref{mc-results} we check the validity of both 
the RS and the $1$-RSB results comparing with Monte Carlo simulations for the internal energy. 
Section~\ref{Renyi-section} and~\ref{I2-section} discuss the classical R\'enyi entropy and the 
mutual information, respectively. Finally, we conclude in section~\ref{conclusions}.

@@

%Clearly, $p_{i_A}\equiv\sum_{j_b\in{\mathcal C}_B}p_{i_A,j_B}$, with $p_{i_A,j_B}$ 
%being the probability that the system is in the configuration $i_A\in{\mathcal C}_A$
%in part $A$, and $j_B\in{\mathcal C}_B$ in part $B$. For a spin model at inverse 
%temperature $\beta$, one has $p_{i_A,j_B}\equiv e^{-\beta E(i_A,j_B)}/Z(\beta)$, 
%where $E(i_A,j_B)$ is the energy of the system, and $Z(\beta)$ its partition function.

Given a generic \emph{classical} spin model at inverse temperature $\beta$, 
the R\'enyi classical entropy $S_n(A)$ is defined as~\cite{jaconis-2013,stephan-2014} 
%
\begin{equation}
\label{renyi}
S_n(A)\equiv \frac{1}{1-n}\log\left(\frac{Z(A,n,\beta)}{Z^n(\beta)}\right).
\end{equation}
%
Here $Z(A,n,\beta)$ is the partition function of the model on the $n$-sheet booklet, 
whereas $Z(\beta)$ is the partition functions on the plane at inverse temperatures $\beta$. 
The booklet geometry is illustrated in Figure~\ref{cartoon}. 
Each sheet is divided into two parts $A$ and $B$ (cf. Figure~\ref{cartoon} (a)). Then $n$ copies 
of the system are ``glued'' together to form the booklet. The spins in part $A$ of the different 
sheets are identified (cf. Figure~\ref{cartoon} (b)). While spins in parts $B$ of the booklet 
are at inverse temperature $\beta$, the ones in $A$ are at the effective temperature $n\beta$. 
A similar geometric construction~\cite{guerra-2002} plays an important role in the mathematical 
proof of the Parisi ansatz.

Notice that one has $S_n(A)\equiv1/(n-1)\log(F(A,n,\beta))-n\log(F(\beta))$, where $F(A,n,\beta)
\equiv\log(Z(A,n\beta))$ and $F(\beta)\equiv\log(Z(\beta))$ are related (apart from a factor 
$\beta$) to the free energy of the model on the booklet and on the plane, respectively.
The mutual information ${\mathcal I}_n(A,B,\beta)$ is defined~\cite{jaconis-2013,stephan-2014} 
%
\begin{equation}
\label{MI}
{\mathcal I}_n(A,B,\beta)\equiv\frac{1}{1-n}\log\Big(
\frac{Z(A,n,\beta)Z(B,n,\beta)}{Z^n(\beta)Z(n\beta)}
\Big),
\end{equation}
%
where $Z(B,n,\beta)$ is obtained from $Z(A,n,\beta)$ by replacing $A$ with $B$. 
Notice that the disorder-averaged mutual information $[{\mathcal I}_n]$ and the R\'enyi 
entropy $[S_n]$ are directly related to the so-called quenched-averaged free energy 
$[F(A,n,\beta)]$, which is the main quantity of interest in disordered systems~\cite{cardy-book}.

For clean (i.e., without disorder) \emph{local} spin models ${\mathcal I}_n$ 
obeys the boundary law 
%
\begin{equation}
{\mathcal I}_n(A,B,\beta)=\alpha_n\ell+{\mathcal G}_n+\gamma_n,
\label{GMI}
\end{equation}
%
with $\ell$ the length of the boundary between $A$ and $B$, and 
$\alpha_n,\gamma_n$ two non-universal constants. Here ${\mathcal G}_n$ 
is the so-called geometric mutual information~\cite{stephan-2014}. 
Interestingly, for critical systems ${\mathcal G}_n$ depends only on the 
geometry of $A$ and $B$, and it is universal. In particular, for conformally 
invariant models ${\mathcal G}_n$ can be calculated using standard methods of 
conformal field theory (CFT), and it allows to numerically 
extract universal information about the CFT, such as the central 
charge~\cite{stephan-2014}.

@@

In this section we present the solution of the S-K model on the booklet, using 
the replica symmetric (RS) approximation. In subsection~\ref{para-section} we focus 
on the high temperature phase, where this approximation is exact, and the 
behavior of the model is fully determined by the diagonal part of the overlap 
tensor $q_{\gamma\gamma'}^{rr'}$. The full structure of $q_{\gamma\gamma'}^{rr'}$ 
in the RS ansatz  is discussed in subsection~\ref{rs-section}. This allows us to 
determine the critical temperature of the paramagnetic-glassy transition.
